Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-01-17 Manuscript/manuscriptPS.tex
Separate systems with additional counterions to other table and move it to SI. The previous table II became too large

\begin{table*}[!p]

@@

\caption{The list of POPC:POPS mixtures simulated with different amounts of added ions. 

@@

\begin{tabular}{l c c c c c c c c c c}
  lipid/counter-ions & force field for lipids / ions & \footnote{Excess Na$^+$ or K$^+$ concentration}C$_{\rm ci}$ (M) & [CaCl$_2$]\,(M)  &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\

@@

%    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 0  & 110:22 & 4935 & 0  & 298  & 100 & 100 \todoi{Equilibration?} & \cite{charmm36pops+83popcT298K}  \\
    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 0 & 250:50 & ?     & 0  & 298  & 200 & ?   & \cite{??} \todoi{Trajectories and further details to be added by J. Madsen}  \\
    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 0 & 110:22 & 4620  & 0  & 298  & 500 & 100 & \cite{charmm36pops+83popcT298Kpiggot}  \\
    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0.45 & 0 & 110:22 & 4926  & 40 & 298  & 200 & 150 & \cite{charmm36pops+83popcT298Kwith450mMK}  \\
    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0.89 & 0 & 110:22 & 4946  & 79 & 298  & 200 & 150 & \cite{charmm36pops+83popcT298Kwith890mMK}  \\
    POPC:POPS (5:1)/Na$^+$ & CHARMM36 \cite{klauda10,venable13} &0 & 0 & 110:22 & 4620  & 0  & 298  & 500 & 100 & \cite{charmm36pops+83popcT298KpiggotSODIUM}  \\
    POPC:POPS (5:1)/Na$^+$  & CHARMM36 \cite{klauda10,venable13} &0.44 & 0 & 110:22 & 4965  & 39 & 298  & 200 & 150 & \cite{charmm36pops+83popcT298Kwith440mMNa}  \\
    POPC:POPS (5:1)/Na$^+$  & CHARMM36 \cite{klauda10,venable13} &0.89 & 0 & 110:22 & 4932  & 79 & 298  & 200 & 150 & \cite{charmm36pops+83popcT298Kwith890mMNa}  \\
    POPC:POPS (5:1)        & CHARMM36 \cite{klauda10,venable13,kim16}  &0 & 0.26 & 250:50 & 11190  & 53  & 298  & 200 & ? \todoi{Equilibration time to be checked}  & \cite{POPC5POPS1withCaClCHARMM}   \\
    POPC:POPS (5:1)        & CHARMM36 \cite{klauda10,venable13,kim16}  &0 & 1 \todoi{Concentration to be checked after the amount of water molecules is known.} & 250:50 & ?  & ?  & 298  & 200 & ?  & \cite{??} \todoi{Trajectories and further details to be added by J. Madsen}  \\
    POPC:POPS (1:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 0 & 150:150 & ?    & 0  & 298  & 200 & ?   & \cite{??} \todoi{Trajectories and further details to be added by J. Madsen}  \\

@@

    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 0   & 120:24 & 5760 & 0    & 298  & 400 & 250 & \cite{POPCpopsMACROG}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0.50  & 0   & 120:24 & 5760 & 52   & 298  & 300 & 200 & \cite{POPCpopsMACROGwithK}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &1.00    & 0   & 120:24 & 5760 & 104  & 298  & 300 & 200 & \cite{POPCpopsMACROGwithK}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &2.00    & 0   & 120:24 & 5760 & 208  & 298  & 300 & 200 & \cite{POPCpopsMACROGwithK}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &3.00    & 0   & 120:24 & 5760 & 311  & 298  & 300 & 200 & \cite{POPCpopsMACROGwithK}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 0.10 & 120:24 & 5760 & 10   & 298  & 600 & 300 & \cite{POPCpopsMACROG}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 0.30 & 120:24 & 5760 & 31   & 298  & 600 & 300  & \cite{POPCpopsMACROG}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 1.00   & 120:24 & 5760 & 104  & 298  & 600 & 300  & \cite{POPCpopsMACROG}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 3.00   & 120:24 & 5760 & 311  & 298  & 600 & 300  & \cite{POPCpopsMACROG}  \\

@@

    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &0  & 0     & 120:24 & 5760 & 0   & 298  & 500 & 200 & \cite{POPCpopsLIPID17withKCI}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &0.5\todoi{Concentration to be checked after the amount of ions is known.}  & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &1\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &2\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &3\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &4\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &0   & 0   & 120:24 & 5760 & 0   & 298  & 500 & 200 & \cite{POPCpopsLIPID17withNaCI}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &0.5\todoi{Concentration to be checked after the amount of ions is known.}  & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &1\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &2\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &3\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &4\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0    & 0      & 60:12 & 3600  & 0   & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0    & 0.08   & 60:12 & 3561  & 5   & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0    & 0.13   & 60:12 & 3561  & 8   & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0    & 0.20   & 60:12 & 3561  & 13  & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0    & 0.41   & 60:12 & 3522  & 26  & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0    & 0.62   & 60:12 & 3483  & 39  & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\

@@

    POPC:POPS (4:1)/Na$^+$  & Berger \cite{tieleman99,mukhopadhyay04} &0    & 0   & 102:26 & 4290 & 0   & 310  & 120 & 80 & \cite{bergerPOPSPOPC4:1mixtureT310K}  \\
    POPC:POPS (4:1)/Na$^+$  & Berger \cite{tieleman99,mukhopadhyay04} &1.03    & 0   & 102:26 & 4290 & 80  & 310  & 200 & 50 & \cite{POPCpopsBERGERwith1000mMNa}  \\
    POPC:POPS (4:1)  & Berger \cite{tieleman99,mukhopadhyay04} &  0  & 0.102\footnote{Calculation of concetration complicated due the scaled ions. Concentration taken as reported in the delivered data.}   & 104:24 & 4306 & 24 & 310  & 300 & 100 & \cite{POPCpopsBERGERwith102mMCa}  \\
    POPC:POPS (4:1)  & Berger \cite{tieleman99,mukhopadhyay04} &  0  & 0.715\footnote{Calculation of concetration complicated due the scaled ions. Concentration taken as reported in the delivered data.}  & 104:24 & 4306 & 72 & 310  & 300 & 100 & \cite{POPCpopsBERGERwith715mMCa}  \\

@@

    POPC:POPS (5:1)/Na$^+$  & GROMOS-CKP \cite{??}             &0    & 0     & 110:22 & ?     & 0  & 298  & 500 & 100 & \cite{POPCpopsGROMOSCKPwithNa}  \\
    POPC:POPS (5:1)/Na$^+$  & GROMOS-CKPM \cite{??}            &0    & 0     & 110:22 & ?     & 0  & 298  & 500 & 100 & \cite{POPCpopsGROMOSCKPMwithNa}  \\

>>>>

\begin{table*}[tb]

@@

\caption{The list of POPC:POPS mixtures simulated with different molar fractions and different amounts of added calcium. 

@@

\begin{tabular}{l c c c c c c c c c }
  lipid/counter-ions & force field for lipids / ions & [CaCl$_2$]\,(M)  &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\

@@

%    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0  & 110:22 & 4935 & 0  & 298  & 100 & 100 \todoi{Equilibration?} & \cite{charmm36pops+83popcT298K}  \\
    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 250:50 & ?     & 0  & 298  & 200 & ?   & \cite{??} \todoi{Trajectories and further details to be added by J. Madsen}  \\
    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 110:22 & 4620  & 0  & 298  & 500 & 100 & \cite{charmm36pops+83popcT298Kpiggot}  \\
    POPC:POPS (5:1)/Na$^+$ & CHARMM36 \cite{klauda10,venable13} &0 & 110:22 & 4620  & 0  & 298  & 500 & 100 & \cite{charmm36pops+83popcT298KpiggotSODIUM}  \\
    POPC:POPS (5:1)        & CHARMM36 \cite{klauda10,venable13,kim16}  & 0.26 & 250:50 & 11190  & 53  & 298  & 200 & ? \todoi{Equilibration time to be checked}  & \cite{POPC5POPS1withCaClCHARMM}   \\
    POPC:POPS (5:1)        & CHARMM36 \cite{klauda10,venable13,kim16}  & 1 \todoi{Concentration to be checked after the amount of water molecules is known.} & 250:50 & ?  & ?  & 298  & 200 & ?  & \cite{??} \todoi{Trajectories and further details to be added by J. Madsen}  \\
    POPC:POPS (1:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 150:150 & ?    & 0  & 298  & 200 & ?   & \cite{??} \todoi{Trajectories and further details to be added by J. Madsen}  \\

@@

    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 120:24 & 5760 & 0    & 298  & 400 & 250 & \cite{POPCpopsMACROG}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} & 0.10 & 120:24 & 5760 & 10   & 298  & 600 & 300 & \cite{POPCpopsMACROG}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} & 0.30 & 120:24 & 5760 & 31   & 298  & 600 & 300  & \cite{POPCpopsMACROG}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} & 1.00   & 120:24 & 5760 & 104  & 298  & 600 & 300  & \cite{POPCpopsMACROG}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} & 3.00   & 120:24 & 5760 & 311  & 298  & 600 & 300  & \cite{POPCpopsMACROG}  \\

@@

    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18}  & 0      & 120:24 & 5760 & 0   & 298  & 500 & 200 & \cite{POPCpopsLIPID17withKCI}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} & 0      & 120:24 & 5760 & 0   & 298  & 500 & 200 & \cite{POPCpopsLIPID17withNaCI}  \\
    POPC:POPS (5:1)         & Lipid14/17 \cite{dickson14,gould18} &0.50    & 120:24 & 5760 & 52   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withCaCl}  \\
    POPC:POPS (5:1)  & Lipid14/17 \cite{dickson14,gould18}        &1.00    & 120:24 & 5760 & 104   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withCaCl}  \\
    POPC:POPS (5:1)  & Lipid14/17 \cite{dickson14,gould18}        &2.00    & 120:24 & 5760 & 208   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withCaCl}  \\
    POPC:POPS (5:1)  & Lipid14/17 \cite{dickson14,gould18}        &3.00    & 120:24 & 5760 & 311   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withCaCl}  \\
    POPC:POPS (5:1)  & Lipid14/17 \cite{dickson14,gould18}        &4.00    & 120:24 & 5760 & 415   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withCaCl}  \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18}  & 0      & 60:12 & 3600  & 0   & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18,smith94,dang06}  & 0.08   & 60:12 & 3561  & 5   & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18,smith94,dang06}  & 0.13   & 60:12 & 3561  & 8   & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18,smith94,dang06}  & 0.20   & 60:12 & 3561  & 13  & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18,smith94,dang06}  & 0.41   & 60:12 & 3522  & 26  & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18,smith94,dang06}  & 0.62   & 60:12 & 3483  & 39  & 298  & 1000  & 1000  & \cite{lipid17_cacl_series}   \\

@@

    POPC:POPS (4:1)/Na$^+$  & Berger \cite{tieleman99,mukhopadhyay04}  & 0   & 102:26 & 4290 & 0   & 310  & 120 & 80 & \cite{bergerPOPSPOPC4:1mixtureT310K}  \\
    POPC:POPS (4:1)  & Berger \cite{tieleman99,mukhopadhyay04}         & 0.102\footnote{Calculation of concetration complicated due the scaled ions. Concentration taken as reported in the delivered data.}   & 104:24 & 4306 & 24 & 310  & 300 & 100 & \cite{POPCpopsBERGERwith102mMCa}  \\
    POPC:POPS (4:1)  & Berger \cite{tieleman99,mukhopadhyay04}         & 0.715\footnote{Calculation of concetration complicated due the scaled ions. Concentration taken as reported in the delivered data.}  & 104:24 & 4306 & 72 & 310  & 300 & 100 & \cite{POPCpopsBERGERwith715mMCa}  \\

@@

    POPC:POPS (5:1)/Na$^+$  & GROMOS-CKP \cite{??}             & 0     & 110:22 & ?     & 0  & 298  & 500 & 100 & \cite{POPCpopsGROMOSCKPwithNa}  \\
    POPC:POPS (5:1)/Na$^+$  & GROMOS-CKPM \cite{??}            & 0     & 110:22 & ?     & 0  & 298  & 500 & 100 & \cite{POPCpopsGROMOSCKPMwithNa}  \\

### 2019-01-17 Manuscript/manuscriptPSsuppl.tex
Separate systems with additional counterions to other table and move it to SI. The previous table II became too large

\usepackage{upgreek}				
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xr}
\externaldocument{manuscriptPS}
\usepackage{soul} % for highlights with \hl{}

@@

@@

The initial structure of the bilayers with 120 POPC and 24 POPS lipids were obtained from CHARMM-GUI~\cite{lee16,jo18}. Bilayers were solvated with 5760 TIP3P water molecules~\cite{jorgensen1983comparison}. For the simulations of the lipids and ions, the Amber Lipid 17 and ff99SB force fields were used, respectively. The simulation steps are the same as the the simulations of DOPS and POPS bilayers described above, with the only difference that the total simulation time was 300~ns, of which last 200~ns was used for the analysis. The number of ions in each specific system is given in Table~\ref{mixedIONsystems}. The trajectory files can be found in Refs~\citenum{POPCpopsLIPID17withKCI,POPCpopsLIPID17withK,POPCpopsLIPID17withNaCI,POPCpopsLIPID17withNa}.\\

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium} \\
The data in Table \ref{mixedIONsystems} and Refs \citenum{lipid17_cacl_series}
\todo{J. Melcr, please write simulation details} \\

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/figS7PC.png}
  \caption{\label{dihedralsHGpc}
    Dihedral angle distributions of the headgroup regions from CHARMM36 and Slipids simulations
    compared between the POPC and POPS lipids. 
    The CHARMM36 POPC simulation is from Ref. \citenum{POPCcharmm36T303K} and Slipids POPC from Ref. \citenum{POPCslipids298K}.
  }
\end{figure*}

\begin{figure*}[]
  \centering
  \includegraphics[width=15.0cm]{../Figs/figS6PC.png}
  \caption{\label{dihedralsGLYpc}
    Dihedral angle distributions of the glycerol backbone regions from CHARMM36 and Slipids simulations
    compared between the POPC and POPS lipids.
    The CHARMM36 POPC simulation is from Ref. \citenum{POPCcharmm36T303K} and Slipids POPC from Ref. \citenum{POPCslipids298K}.
  }
\end{figure*}

>>>>

\usepackage{rotating}
\usepackage{upgreek}				
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xr}
\externaldocument{manuscriptPS}
\usepackage{soul} % for highlights with \hl{}

@@

%\begin{table*}[!p]
\begin{sidewaystable*}[!p]
\centering
\caption{The list of POPC:POPS mixtures simulated with different amounts of added monovalent ions. 
  The salt concentrations are calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
  This corresponds the concentration in buffer before solvating lipids, which are
  reported in the experiments by Roux et al.~\cite{roux90}.
  The simulation details are given in the supplementary information.
   % The lipid force fields named as in our previous work~\cite{botan15}.
}\label{mixedIONsystemsMONOVALENT}
%begin{minipage}[t]{\textwidth}
\begin{tabular}{l c c c c c c c c c}
  lipid/counter-ions & force field for lipids / ions & \footnote{Excess Na$^+$ or K$^+$ concentration}C$_{\rm ci}$ (M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
  \hline
%    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 0  & 110:22 & 4935 & 0  & 298  & 100 & 100 \todoi{Equilibration?} & \cite{charmm36pops+83popcT298K}  \\
    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 250:50 & ?     & 0  & 298  & 200 & ?   & \cite{??} \todo{Trajectories and further details to be added by J. Madsen}  \\
    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0     & 110:22 & 4620  & 0  & 298  & 500 & 100 & \citenum{charmm36pops+83popcT298Kpiggot}  \\
    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0.45  & 110:22 & 4926  & 40 & 298  & 200 & 150 & \citenum{charmm36pops+83popcT298Kwith450mMK}  \\
    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0.89  & 110:22 & 4946  & 79 & 298  & 200 & 150 & \citenum{charmm36pops+83popcT298Kwith890mMK}  \\
    POPC:POPS (5:1)/Na$^+$ & CHARMM36 \cite{klauda10,venable13} &0     & 110:22 & 4620  & 0  & 298  & 500 & 100 & \citenum{charmm36pops+83popcT298KpiggotSODIUM}  \\
    POPC:POPS (5:1)/Na$^+$  & CHARMM36 \cite{klauda10,venable13} &0.44 & 110:22 & 4965  & 39 & 298  & 200 & 150 & \citenum{charmm36pops+83popcT298Kwith440mMNa}  \\
    POPC:POPS (5:1)/Na$^+$  & CHARMM36 \cite{klauda10,venable13} &0.89 & 110:22 & 4932  & 79 & 298  & 200 & 150 & \citenum{charmm36pops+83popcT298Kwith890mMNa}  \\
    \hline
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 120:24 & 5760 & 0    & 298  & 400 & 250 & \citenum{POPCpopsMACROG}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0.50 & 120:24 & 5760 & 52   & 298  & 300 & 200 & \citenum{POPCpopsMACROGwithK}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &1.00 & 120:24 & 5760 & 104  & 298  & 300 & 200 & \citenum{POPCpopsMACROGwithK}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &2.00 & 120:24 & 5760 & 208  & 298  & 300 & 200 & \citenum{POPCpopsMACROGwithK}  \\
    POPC:POPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &3.00 & 120:24 & 5760 & 311  & 298  & 300 & 200 & \citenum{POPCpopsMACROGwithK}  \\
    \hline
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &0     & 120:24 & 5760 & 0   & 298  & 500 & 200 & \citenum{POPCpopsLIPID17withKCI}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &0.50  & 120:24 & 5760 & 54   & 298  & 300 & 200 & \citenum{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &1.00  & 120:24 & 5760 & 104  & 298  & 300 & 200 & \citenum{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &2.00  & 120:24 & 5760 & 208  & 298  & 300 & 200 & \citenum{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &3.00  & 120:24 & 5760 & 311  & 298  & 300 & 200 & \citenum{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &4.00  & 120:24 & 5760 & 415  & 298  & 300 & 200 & \citenum{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &0    & 120:24 & 5760 & 0   & 298  & 500 & 200 & \citenum{POPCpopsLIPID17withNaCI}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &0.50 & 120:24 & 5760 & 54   & 298  & 300 & 200 & \citenum{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &1.00 & 120:24 & 5760 & 104   & 298  & 300 & 200 & \citenum{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &2.00 & 120:24 & 5760 & 208   & 298  & 300 & 200 & \citenum{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &3.00 & 120:24 & 5760 & 311   & 298  & 300 & 200 & \citenum{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &4.00 & 120:24 & 5760 & 415   & 298  & 300 & 200 & \citenum{POPCpopsLIPID17withNa}  \\
    \hline
    POPC:POPS (4:1)/Na$^+$  & Berger \cite{tieleman99,mukhopadhyay04} &0    & 102:26 & 4290 & 0   & 310  & 120 & 80 & \citenum{bergerPOPSPOPC4:1mixtureT310K}  \\
    POPC:POPS (4:1)/Na$^+$  & Berger \cite{tieleman99,mukhopadhyay04} &1.03 & 102:26 & 4290 & 80  & 310  & 200 & 50 & \citenum{POPCpopsBERGERwith1000mMNa}  \\
\end{tabular}
%\end{minipage}
\end{sidewaystable*}

% \end{table*}

@@

The initial structure of the bilayers with 120 POPC and 24 POPS lipids were obtained from CHARMM-GUI~\cite{lee16,jo18}. Bilayers were solvated with 5760 TIP3P water molecules~\cite{jorgensen1983comparison}. For the simulations of the lipids and ions, the Amber Lipid 17 and ff99SB force fields were used, respectively. The simulation steps are the same as the the simulations of DOPS and POPS bilayers described above, with the only difference that the total simulation time was 300~ns, of which last 200~ns was used for the analysis. The number of ions in each specific system is given in Table~\ref{mixedIONsystemsMONOVALENT}. The trajectory files can be found in Refs~\citenum{POPCpopsLIPID17withKCI,POPCpopsLIPID17withK,POPCpopsLIPID17withNaCI,POPCpopsLIPID17withNa}.\\

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium} \\
The data in Table \ref{mixedIONsystems} and Refs \citenum{lipid17_cacl_series,POPCpopsLIPID17withCaCl} 
\todo{B. Kav and J. Melcr, please write simulation details} \\

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/figS7PC.png}
  \caption{\label{dihedralsHGpc}
    Dihedral angle distributions of the headgroup regions from CHARMM36 and Slipids simulations
    compared between the POPC and POPS lipids. 
    The CHARMM36 POPC simulation is from Ref. \citenum{POPCcharmm36T303K} and Slipids POPC from Ref. \citenum{POPCslipids298K}.
  }
  \todo{Colours with more contrast would be better. The y-axis scale could be zoomed.}
\end{figure*}

\begin{figure*}[]
  \centering
  \includegraphics[width=15.0cm]{../Figs/figS6PC.png}
  \caption{\label{dihedralsGLYpc}
    Dihedral angle distributions of the glycerol backbone regions from CHARMM36 and Slipids simulations
    compared between the POPC and POPS lipids.
    The CHARMM36 POPC simulation is from Ref. \citenum{POPCcharmm36T303K} and Slipids POPC from Ref. \citenum{POPCslipids298K}.
  }
  \todo{Colours with more contrast would be better. The y-axis scale could be zoomed.}
\end{figure*}
